Do investors in clean energy ETFs herd? The role of climate risks

### 2025-09-24 alt_energy_herding_corrected.tex
first review comments

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_files/figure-pdf/fig-csad-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_files/figure-pdf/fig-rolling_window-1.pdf}}

>>>>

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_corrected_files/figure-pdf/fig-csad-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_corrected_files/figure-pdf/fig-rolling_window-1.pdf}}

### 2025-09-24 alt_energy_herding_final.qmd
first review comments

The results from the Probit model are reported in @tbl-probit_analysis. Showing that only the transitional climate risk index significantly decreases the probability of herding in clean energy ETFs.^[It should be noted that due to the availability of climate risk data from @bua2024, the probit analysis covers the period from May 1, 2016 to December 30, 2023.] Transitional climate risk represent good news for investors in clean energy stocks, possibly reducing their self-reinforcing nature of confidence in the tyranny of the majority [@teraji2003]) and their need for shared intention and action, resulting in a decrease in the herding behaviour. In the presence of higher transitional risk with respect to the climate, clean energy ETFs become a more attractive investment alternative for environmentally conscious investors who allocate their money to alternative energy investment products [see @bouri2022], reinforcing the confidence of investors in their own information. As a result, the cross-sectional dispersion of clean energy ETFs tends to increase. Our results are somewhat in line with @bouri2023 who show that the transitional climate risk is more important than physical risk for the return and volatility of clean energy stocks. They also concord with other relevant studies which indicate that in the event of climate policy shocks, clean energy assets could serve the role of hedging instruments [@gabriel2024] and tend to outperform brown energy assets [@bouri2022].

>>>>

The results from the Probit model are reported in @tbl-probit_analysis. Showing that only the transitional climate risk index significantly decreases the probability of herding in clean energy ETFs.^[It should be noted that due to the availability of climate risk data from @bua2024, the probit analysis covers the period from May 1, 2016 to December 30, 2023.] Transitional climate risk represent good news for investors in clean energy stocks, possibly reducing their self-reinforcing nature of confidence in the tyranny of the majority [@teraji2003] and their need for shared intention and action, resulting in a decrease in the herding behaviour. In the presence of higher transitional risk with respect to the climate, clean energy ETFs become a more attractive investment alternative for environmentally conscious investors who allocate their money to alternative energy investment products [see @bouri2022], reinforcing the confidence of investors in their own information. As a result, the cross-sectional dispersion of clean energy ETFs tends to increase. Our results are somewhat in line with @bouri2023 who show that the transitional climate risk is more important than physical risk for the return and volatility of clean energy stocks. They also concord with other relevant studies which indicate that in the event of climate policy shocks, clean energy assets could serve the role of hedging instruments [@gabriel2024] and tend to outperform brown energy assets [@bouri2022].

### 2025-09-24 alt_energy_herding_final.tex
first review comments

This study offers new insights into the herding behaviour in US clean energy Exchange-traded funds (ETFs) over the period from May 1, 2016 to June 19, 2024. 
The baseline model shows significant herding. An extended mode indicates that herding is present in both down and up markets, with a stronger effect in the down market, suggesting an asymmetry. 
Herding is also found to be time-varying. Further analysis shows that the transition climate risk, particularly its high level, reduces the probability of herding in clean energy ETFs, whereas physical climate risk does not exert any significant impact on the probability of herding. 
Thus, large levels of transition climate risk seem to encourage market efficiency in clean energy ETFs  and climate-hedging behaviour by investors.

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_corrected_files/figure-pdf/fig-csad-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_corrected_files/figure-pdf/fig-rolling_window-1.pdf}}

@@

The results from the Probit model are reported in
Table~\ref{tbl-probit_analysis}. Showing that only the transitional
climate risk index significantly decreases the probability of herding in
clean energy ETFs.\footnote{It should be noted that due to the
  availability of climate risk data from \citet{bua2024}, the probit
  analysis covers the period from May 1, 2016 to December 30, 2023.}
Transitional climate risk represent good news for investors in clean
energy stocks, possibly reducing their self-reinforcing nature of
confidence in the tyranny of the majority \citep{teraji2003}) and their
need for shared intention and action, resulting in a decrease in the
herding behaviour. In the presence of higher transitional risk with
respect to the climate, clean energy ETFs become a more attractive
investment alternative for environmentally conscious investors who
allocate their money to alternative energy investment products
\citep[see][]{bouri2022}, reinforcing the confidence of investors in
their own information. As a result, the cross-sectional dispersion of
clean energy ETFs tends to increase. Our results are somewhat in line
with \citet{bouri2023} who show that the transitional climate risk is
more important than physical risk for the return and volatility of clean
energy stocks. They also concord with other relevant studies which
indicate that in the event of climate policy shocks, clean energy assets
could serve the role of hedging instruments \citep{gabriel2024} and tend
to outperform brown energy assets \citep{bouri2022}.

>>>>

\textbf{Purpose} -- This study investigates herding behaviour in US clean energy Exchange-traded funds (ETFs) and examines the role of climate risks in influencing such behaviour over the period from May 1, 2016 to June 19, 2024.

\textbf{Design/methodology/approach} -- We employ a baseline herding model and extend it to examine asymmetric effects across market conditions. The analysis incorporates time-varying herding measures and examines the impact of both transitional and physical climate risks on herding probability using regression techniques.

\textbf{Findings} -- The baseline model reveals significant herding behaviour in clean energy ETFs. The extended model indicates that herding is present in both down and up markets, with a stronger effect in down markets, suggesting asymmetry. Herding is also found to be time-varying. Notably, high levels of transitional climate risk reduce the probability of herding in clean energy ETFs, whereas physical climate risk does not exert any significant impact on herding probability.

\textbf{Research limitations/implications} -- The study focuses specifically on US clean energy ETFs over a defined period, which may limit generalizability to other markets or asset classes. The findings provide insights into the behavioral dynamics of sustainable investment markets during periods of varying climate risk.

\textbf{Practical implications} -- The results suggest that high levels of transitional climate risk encourage market efficiency in clean energy ETFs and promote climate-hedging behaviour by investors. This has important implications for portfolio managers and policymakers in understanding market dynamics in sustainable finance.

\textbf{Originality/value} -- This study provides novel empirical evidence on the relationship between climate risks and herding behaviour in clean energy ETFs, contributing to the growing literature on behavioral finance in sustainable investment markets.

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_final_files/figure-pdf/fig-csad-1.pdf}}

@@

\pandocbounded{\includegraphics[keepaspectratio]{alt_energy_herding_final_files/figure-pdf/fig-rolling_window-1.pdf}}

@@

The results from the Probit model are reported in
Table~\ref{tbl-probit_analysis}. Showing that only the transitional
climate risk index significantly decreases the probability of herding in
clean energy ETFs.\footnote{It should be noted that due to the
  availability of climate risk data from \citet{bua2024}, the probit
  analysis covers the period from May 1, 2016 to December 30, 2023.}
Transitional climate risk represent good news for investors in clean
energy stocks, possibly reducing their self-reinforcing nature of
confidence in the tyranny of the majority \citep{teraji2003} and their
need for shared intention and action, resulting in a decrease in the
herding behaviour. In the presence of higher transitional risk with
respect to the climate, clean energy ETFs become a more attractive
investment alternative for environmentally conscious investors who
allocate their money to alternative energy investment products
\citep[see][]{bouri2022}, reinforcing the confidence of investors in
their own information. As a result, the cross-sectional dispersion of
clean energy ETFs tends to increase. Our results are somewhat in line
with \citet{bouri2023} who show that the transitional climate risk is
more important than physical risk for the return and volatility of clean
energy stocks. They also concord with other relevant studies which
indicate that in the event of climate policy shocks, clean energy assets
could serve the role of hedging instruments \citep{gabriel2024} and tend
to outperform brown energy assets \citep{bouri2022}.

### 2025-09-24 title.tex
second journal submission
first review comments

ipsum

>>>>

\textbf{Purpose} -- This study investigates herding behaviour in US clean energy Exchange-traded funds (ETFs) and examines the role of climate risks in influencing such behaviour over the period from May 1, 2016 to June 19, 2024.

\textbf{Design/methodology/approach} -- We employ a baseline herding model and extend it to examine asymmetric effects across market conditions. The analysis incorporates time-varying herding measures and examines the impact of both transitional and physical climate risks on herding probability using regression techniques.

\textbf{Findings} -- The baseline model reveals significant herding behaviour in clean energy ETFs. The extended model indicates that herding is present in both down and up markets, with a stronger effect in down markets, suggesting asymmetry. Herding is also found to be time-varying. Notably, high levels of transitional climate risk reduce the probability of herding in clean energy ETFs, whereas physical climate risk does not exert any significant impact on herding probability.

\textbf{Research limitations/implications} -- The study focuses specifically on US clean energy ETFs over a defined period, which may limit generalizability to other markets or asset classes. The findings provide insights into the behavioral dynamics of sustainable investment markets during periods of varying climate risk.

\textbf{Practical implications} -- The results suggest that high levels of transitional climate risk encourage market efficiency in clean energy ETFs and promote climate-hedging behaviour by investors. This has important implications for portfolio managers and policymakers in understanding market dynamics in sustainable finance.

\textbf{Originality/value} -- This study provides novel empirical evidence on the relationship between climate risks and herding behaviour in clean energy ETFs, contributing to the growing literature on behavioral finance in sustainable investment markets.
